---
title: "HW1"
author: "Debi Gonzalez"
date: "2025-09-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(swirl)
```


For example...
###
> library("swirl")
| Hi! Type swirl() when you are ready to begin.
> swirl()
| Welcome to swirl! Please sign in. If you've been here
| before, use the same name as you did then. If you are new,
| call yourself something unique.
What shall I call you?
| Would you like to continue with one of these lessons?
1: R Programming Basic Building Blocks
2: No. Let me start something new.
Selection: 1
### Work through the first five lessons (Basic Building Blocks, Workspace and Files, Sequences of Numbers, Vectors, Missing Values). Then answer the following questions. Show work (copy the R commands) for each question that requires output from R.


1. Please define a vector of the first ten even numbers after zero (e.g., 2, 4, 6...). First print it to the screen, and then save it in a variable of your choice. Take the square root of that variable and save the result to a new variable name of your choice.

```{r}
c(2,4,6,8,10,12,14,16,18,20)

x <- c(2,4,6,8,10,12,14,16,18,20)

sqrt(x)

y <- sqrt(x)
```


2. Find your directory and list all objects in the directory.
```{r}
getwd()
ls()
```


3. Print a sequence of 50 numbers, incremented by .7 at each step. Use a simple command to show the length of the sequence.
```{r}
my_seq <- seq(0, by=0.7, length=50)

print(my_seq)

length(my_seq)
```



4. Define the logical values of |, &, and !A.

`|` = OR  
`&` = AND  
`!A` = NOT (or negation)

*These are logical operators used to combine or negate logical expressions, which result in TRUE or FALSE values.*



5. How do you tell R that you are dealing with a string variable?

*Typically, I'd tell R that I'm dealing with a string variable by enclosing the text in quotes (i.e. "x").*



6. How do you represent a value that is Not Available or Missing in R?

*Typically, Not Available or Missing values are represented by NA.*



7. Create a vector of 250 draws from a standard normal distribution.
```{r}
my_vector <- rnorm(250)
my_vector
```



8. Create the following vector: 1,4,3,4,2,NA,5,NA,4,5,2,NA. Assign the vector to a variable of your choice. Pull the NA values from that vector into a new vector and output that vector.

```{r}
v <- c(1,4,3,4,2,NA,5,NA,4,5,2,NA)
v_na <- v[is.na(v)]
v_na
```


