---
title: "HW3"
author: "Debi Gonzalez"
date: "2025-09-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Work through the Logic lesson, and the first half of the Functions lesson (go to 50%). Then answer the following questions. Show work (copy the R commands) for each question that requires output from R.

1.	List all of the logical operators you learned in the “Logic” section and define each.


2.	Find the opposite of whether 3 == 4.

```{r}

# Option 1
3 != 4

# Option 2
!(3 == 4)
```


3.	Evaluate !(7 > 6) & (7==7) | (6==5)
```{r, echo=True}
!(7 > 6) & (7==7) | (6==5)
# First statement is false, it's the negation of the TRUE statement in paranthesis.
# second statement is true
# Third statemenet is false. 
```


4.	Which is evaluated first for each pair below?

  a.	&  |
  b.	&&  |
  c.	&& ||
  d.	& ||

5.	What does the xor() function do?
In R, the xor() function performs a logical exclusive OR operation. It takes two logical vectors or individual logical values as input and returns TRUE if exactly one of the inputs is TRUE, and FALSE otherwise.

6.	Tell R to tell you the date.
```{r, echo=True}
# getting today's date
Sys.Date()
```


7.	Calculate the mean of 3 numbers using a simple R function that comes with R.
```{r}

# To make this exercise fun, I'm creating a vector of 3 numbers randomly selected between 1 and 100. Then, I'm printing my vector to see what we are working with and using the mean() function.

vector <- sample(1:100, 3, replace = TRUE)
print(vector)
mean(vector)
```


8.	Print the output of: boring_function('My first function!')

```{r}

boring_function <- function(x) {x}
boring_function('My first function!')


```
```{r}

# Playing around with functions for my own practice.

example_func <- function(y, z) {
  print(y)
  print(z)
  y/z
}

example_func(y=2, z=4)
```


9.	Write the function my_mean from the tutorial and print your function here.

```{r}
my_mean <- function(x) {sum(x)/length(x)}

values <- c(1,4,5,3,4,5)

my_mean(values)
```

10.	After writing your function, provide the output of the following:
  
  a.	remainder(4, div = 2)

  b.	remainder(4, div = 10)
