---
title: "HW3"
author: "Debi Gonzalez"
date: "2025-09-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Work through the Logic lesson, and the first half of the Functions lesson (go to 50%). Then answer the following questions. Show work (copy the R commands) for each question that requires output from R.

1.	List all of the logical operators you learned in the “Logic” section and define each.

These return logical values like TRUE or FALSE.
== Equal to – checks if two values are exactly equal.
!= Not equal to – checks if two values are different.
< Less than.
<= Less than or equal to.
> Greater than.
>= Greater than or equal to.

Logical Operators
! Logical NOT - negates a logical value
& Element-wise AND; it's used to evaluate AND across a vector.
&& Short-circuit AND; it evaluates only the first element of a vector and returns a single TRUE or FALSE.
| Element-wise OR; it's used to evaluate OR across a vector.
|| Short-circuit OR; it evaluates only the first element of a vector and returns a single TRUE or FALSE.

2.	Find the opposite of whether 3 == 4.

```{r}

# Option 1
3 != 4

# Option 2
!(3 == 4)
```


3.	Evaluate !(7 > 6) & (7==7) | (6==5)
```{r, echo=True}
!(7 > 6) & (7==7) | (6==5)
# First statement is false, it's the negation of the TRUE statement in paranthesis.
# second statement is true
# Third statement is false. 
# Final statement is ...
```


4.	Which is evaluated first for each pair below?

  a.	&  | --> & is evaluated first
  b.	&&  | --> | is evaluated first
  c.	&& || --> && is evaluated first
  d.	& || --> & is evaluate first
  
  
```{r}
# My reference hierarchy based on ?syntax

&    (element-wise AND)
|    (element-wise OR)
&&   (short-circuit AND)
||   (short-circuit OR)

```
  

5.	What does the xor() function do?
In R, the xor() function performs a logical exclusive OR operation. It takes two logical vectors or individual logical values as input and returns TRUE if exactly one of the inputs is TRUE, and FALSE otherwise.

6.	Tell R to tell you the date.
```{r, echo=True}
# getting today's date
Sys.Date()
```


7.	Calculate the mean of 3 numbers using a simple R function that comes with R.
```{r}

# To make this exercise fun, I'm creating a vector of 3 numbers randomly selected between 1 and 100. Then, I'm printing my vector to see what we are working with and using the mean() function.

vector <- sample(1:100, 3, replace = TRUE)
print(vector)
mean(vector)
```


8.	Print the output of: boring_function('My first function!')

```{r}

boring_function <- function(x) {x}
boring_function('My first function!')


```


```{r}

# Playing around with functions for my own practice.

example_func <- function(y, z) {
  print(y)
  print(z)
  y/z
}

example_func(y=2, z=4)
```


9.	Write the function my_mean from the tutorial and print your function here.

```{r}
my_mean <- function(x) {sum(x)/length(x)}

values <- c(1,4,5,3,4,5)

my_mean(values)
```

10.	After writing your function, provide the output of the following:

```{r}
remainder <- function(num, divisor = 2) {
  num %% divisor 
  
}
```

  
  a.	remainder(4, div = 2)
```{r}
remainder(4)

# We can just put 4 as num since default divisor is 2.
```
  

  b.	remainder(4, div = 10)
```{r} 

# We must include both num and divisor bc our divisor is now different from default.
remainder(4,10)
```

